---
title: "Assignment3"
author: "RF"
date: "3/6/2018"
output: html_document
---

```{r setup, include=FALSE}
setwd("~/Library/Mobile Documents/com~apple~CloudDocs/4. Aarhus Universitet/4. Semester/1. Computational Modelling/RFOLDER/ASS3_Multivariate linear models/ASS3_Multivariate")

pacman::p_load("coda", "mvtnorm", "devtools", "rstan", "rethinking", "dplyr", "dfcrm", "brms")
pacman::p_load(brms)

# library(devtools)
# devtools::install_github("paul-buerkner/brms")
# pacman::p_load(brms)

install.packages("brms", dependencies = T)
library(brms)
update.packages("brms", dependencies = T)

theme_set(theme_default()) 

#data
data <- read.csv("Assignment3Data.csv")


```

## Multivariate linear models

In this assignment you will investigate the relation between different aspects of IQ and symptom severity in ASD (as measured by ADOS). The assignment is meant to make you practice on linear models, multiple predictors and interactions.

The data you will use is the outcome of the psychological testing of the children you analyzed in methods 3.
Data: https://www.dropbox.com/s/hom2qnmp9hyfwmu/Assignment3Data.csv?dl=0
The data consists of ChildID, gender, age, diagnosis (ASD), symptom severity (ADOS), several aspects of IQ (NonVerbal/Perceptual,Verbal,Social) and the visit at which the tests were administered. The questions should be answered by focusing on visit 1.

The questions you will have to answer are:

### 1. Assess the relation between symptom severity and IQ (focus on visit 1 and children with ASD) and report the model, a plot of the model, a couple of lines describing the quality of the model and interpreting the results.
P.S. Should you scale? 
P.P.S. Don't forget to motivate your priors. 
P.P.P.S. At least one plot for results and a plot for quality of each model (here and in the next questions) would be appreciated.

```{r df}
# only visit 1
v1_asd <- filter(data, data$Visit == 1 & data$ASD == 1)
```


1.1. Verbal IQ and ADOS
```{r vIQ + ados}
plot(v1_asd$ADOS ~ v1_asd$VerbalIQ)  

#  defining model first
m0 <-  bf(ADOS ~ 1 + VerbalIQ)
dens (v1_asd$VerbalIQ)
dens (v1_asd$ADOS)

#Mean centering
v1_asd$ADOS = (v1_asd$ADOS-mean(v1_asd$ADOS, na.rm = T)) / sd(v1_asd$ADOS)
v1_asd$VerbalIQ = (v1_asd$VerbalIQ-mean(v1_asd$VerbalIQ, na.rm = T)) / sd(v1_asd$VerbalIQ)

#Getting the priors
get_prior(m0, v1_asd, family = gaussian)

#Defining priors used
prior = c(prior(normal(0,0.5), class = 'Intercept'), prior(normal(0,1), class = 'b'), prior(normal(0,0.8), class = 'sigma'))

#modelling
m0_prior <-  brm(m0, family = gaussian, v1_asd, prior = prior, sample_prior = "only")

#Checking priors
pp_check(m0_prior, nsample = 100)


#MAP
m.m0 <- map( 
  alist(
    ADOS ~ dnorm(mu, sigma),
    mu ~ a + b * VerbalIQ,
    a ~ dnorm(0,0.5), 
    b ~ dnorm(0,1),
    sigma ~ dunif(0,0.8)
  ),
  data = v1_asd)
precis(m.m0)


### Plot
# extract 20 samples from the posterior
post <- extract.samples( m.m0 , n=100 )
# display raw data and sample size
plot( ADOS ~ VerbalIQ , data=v1_asd )
abline( a=coef(m.m0)["a"] , b=coef(m.m0)["b"], col=col.alpha("red", 1))
for (i in 1:150)
abline( a=post$a[i], b=post$b[i], col=col.alpha("blue",0.1))



### new plot
# define sequence of weights to compute predictions for
# these values will be on the horizontal axis
verb.seq <- seq(from= -2, to= 2.5, by=0.1)
# use link to compute mu
# for each sample from posterior
# and for each weight in weight.seq
mu <- link( m.m0 , data=data.frame(VerbalIQ=verb.seq) )
str(mu)

###FUNKYPLOT
# use type="n" to hide raw data
plot( ADOS~ VerbalIQ , v1_asd , type="n" )
# loop over samples and plot each mu value
for ( i in 1:nrow(mu) )
points( verb.seq , mu[i,] , pch=16 , col=col.alpha(rangi2,0.1) )

### GOODPPLOT NICEPLOT
# summarize the distribution of mu
mu.mean <- apply(mu , 2 , mean ) 
mu.HPDI <- apply(mu , 2 , HPDI , prob=0.89 )
# plot raw data
# fading out points to make line and interval more visible
plot( ADOS ~ VerbalIQ , data=v1_asd , col=col.alpha(rangi2,1) )
# plot the MAP line, aka the mean mu for each weight
lines( verb.seq , mu.mean )
# plot a shaded region for 89% HPDI
shade( mu.HPDI , verb.seq )
```

1.2. Non Verbal IQ and ADOS
```{r nIQ + ados} 
plot(v1_asd$ADOS ~ v1_asd$NonVerbalIQ) 

#  defining model first
m2 <-  bf(ADOS ~ 1 + NonVerbalIQ)
dens (v1_asd$NonVerbalIQ)
dens (v1_asd$ADOS)

#Mean centering
v1_asd$NonVerbalIQ = (v1_asd$NonVerbalIQ-mean(v1_asd$NonVerbalIQ, na.rm = T)) / sd(v1_asd$NonVerbalIQ)

#Getting the priors
get_prior(m2, v1_asd, family = gaussian)

#Defining priors used
prior = c(prior(normal(0,0.5), class = 'Intercept'), prior(normal(0,0.5), class = 'b'), prior(normal(0,0.8), class = 'sigma'))

#modelling
m2_prior <-  brm(m2, family = gaussian, v1_asd, prior = prior, sample_prior = "only")

#Checking priors
pp_check(m0_prior, nsample = 100)


#MAP
m.m2 <- map( 
  alist(
    ADOS ~ dnorm(mu, sigma),
    mu ~ a + b * NonVerbalIQ,
    a ~ dnorm(0,0.5), 
    b ~ dnorm(0,0.5),
    sigma ~ dunif(0,0.08)
  ),
  data = v1_asd)
precis(m.m2)


### Plot
# extract 20 samples from the posterior
post <- extract.samples( m.m2 , n=100 )
# display raw data and sample size
plot( ADOS ~ NonVerbalIQ , data=v1_asd )
abline( a=coef(m.m2)["a"] , b=coef(m.m2)["b"], col=col.alpha("red", 1))
for (i in 1:150)
abline( a=post$a[i], b=post$b[i], col=col.alpha("blue",0.1))



### new plot
# define sequence of weights to compute predictions for
# these values will be on the horizontal axis
nverb.seq <- seq(from= -3, to= 3, by=0.1)
# use link to compute mu
# for each sample from posterior
# and for each weight in weight.seq
mu2 <- link( m.m2 , data=data.frame(NonVerbalIQ=nverb.seq) )
str(mu2)

###FUNKYPLOT 
# use type="n" to hide raw data
plot( ADOS ~ NonVerbalIQ , v1_asd , type="n" )
# loop over samples and plot each mu value
for ( i in 1:nrow(mu2) )
points( nverb.seq , mu2[i,] , pch=16 , col=col.alpha(rangi2,0.1) )

### GOODPPLOT NICEPLOT
# summarize the distribution of mu
mu2.mean <- apply(mu2 , 2 , mean )
mu2.HPDI <- apply(mu2 , 2 , HPDI , prob=0.89 )
# plot raw data
# fading out points to make line and interval more visible
plot( ADOS ~ NonVerbalIQ , data=v1_asd , col=col.alpha(rangi2,1) )
# plot the MAP line, aka the mean mu for each weight
lines( nverb.seq , mu2.mean )
# plot a shaded region for 89% HPDI
shade( mu2.HPDI , nverb.seq )
```

1.3. Social IQ and ADOS
```{r sIQ + ados} 
plot(v1_asd$ADOS ~ v1_asd$SocialIQ)

#  defining model first
m3 <-  bf(ADOS ~ 1 + SocialIQ)
dens (v1_asd$SocialIQ)
dens (v1_asd$ADOS)

#Mean centering
v1_asd$SocialIQ = (v1_asd$SocialIQ-mean(v1_asd$SocialIQ, na.rm = T)) / sd(v1_asd$SocialIQ)

#Getting the priors
get_prior(m3, v1_asd, family = gaussian)

#Defining priors used
prior = c(prior(normal(0,0.5), class = 'Intercept'), prior(normal(0,0.5), class = 'b'), prior(normal(0,0.8), class = 'sigma'))

#modelling
m3_prior <-  brm(m3, family = gaussian, v1_asd, prior = prior, sample_prior = "only")

#Checking priors
pp_check(m3_prior, nsample = 100)


#MAP
m.m3 <- map( 
  alist(
    ADOS ~ dnorm(mu, sigma),
    mu ~ a + b * SocialIQ,
    a ~ dnorm(0,0.5), 
    b ~ dnorm(0,0.5),
    sigma ~ dunif(0,0.8)
  ),
  data = v1_asd)
precis(m.m3)


### Plot
# extract 20 samples from the posterior
post <- extract.samples( m.m3 , n=100 )
# display raw data and sample size
plot( ADOS ~ SocialIQ , data=v1_asd )
abline( a=coef(m.m3)["a"] , b=coef(m.m3)["b"], col=col.alpha("red", 1))
for (i in 1:150)
abline( a=post$a[i], b=post$b[i], col=col.alpha("blue",0.1))



### new plot
# define sequence of weights to compute predictions for
# these values will be on the horizontal axis
soc.seq <- seq(from= -2, to= 3, by=0.1)
# use link to compute mu
# for each sample from posterior
# and for each weight in weight.seq
mu3 <- link( m.m3, data=data.frame(SocialIQ=soc.seq) )
str(m3)

###FUNKYPLOT
# use type="n" to hide raw data
plot( ADOS~ SocialIQ , v1_asd , type="n" )
# loop over samples and plot each mu value
for ( i in 1:nrow(mu3) ) 
points( soc.seq , mu3[i,] , pch=16 , col=col.alpha(rangi2,0.1) )

### GOODPPLOT NICEPLOT
# summarize the distribution of mu
mu3.mean <- apply(mu3 , 2 , mean )
mu3.HPDI <- apply(mu3 , 2 , HPDI , prob=0.89 )
# plot raw data
# fading out points to make line and interval more visible
plot( ADOS ~ SocialIQ , data=v1_asd , col=col.alpha(rangi2,1) )
# plot the MAP line, aka the mean mu for each weight
lines( soc.seq , mu3.mean )
# plot a shaded region for 89% HPDI
shade( mu3.HPDI , soc.seq )
```


### 2. Do the different aspects of IQ account for different portions of the variance in ADOS? 
2.1. Does it make sense to have all IQ measures in the same model? First write a few lines answering the question and motivating your answer, including a discussion as to what happens when you put all of them in the same model. Then build a model following your answer. If your answer is "no", you are not free, you still have to answer: are there alternative ways of answering the question?
```{r}


```

2.2. Build the model, assess its quality, write a few lines interpreting the results.
```{r}

```


### 3. Let's now include also the TD children. Does it make sense to ask whether IQ and ADOS are related? Motivate your answer. In any case, if you wanted to build a model to answer that question, which model would you build? Run the model, assess its quality, write a few lines interpreting the results.


### 4. Let's discuss contents:
4.1. You have three scores for IQ, do they show shared variance? Is that the same in TD and ASD? What does that tell us about IQ?
4.2. You have explored the relation between IQ and ADOS. How do you explain that relation from a cognitive perspective? N.B. You can present alternative hypotheses.




